Virtualisierung eines Echtzeit-Betriebssystems zur Steuerung eines Roboters mit Schwerpunkt auf die Einhaltung der Echtzeit

### 2024-06-07 Masterarbeit_Robotik_Pamuk.tex
[DRAFT] Host Configurations 5/13 finished

\noindent Figure~\ref{fig:lstopo} is the output of the \texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices. 
\begin{figure}[H]
	\centering
	\includegraphics[width=1.0\columnwidth]{img/lstopo.png}
	\caption[Hardware topology]{Hardware topology}
	\label{fig:lstopo}
\end{figure}

@@

\section{Structure}
\noindent Salamander 4 is the proprietary operating system of SIGMATEK. It is based on Linux version 5.15.94 and integrates Xenomai 3.2, a real-time development environment~\cite{XenomaiXenomai}. Salamander 4 is a 64-bit system, which refers to the x86\_64 architecture. The real-time behaviour is achieved through the use of Symmetric Multi-Processing (SMP) and Preemptive Scheduling (PREEMPT). In addition, it uses IRQPIPE to process interrupts in a way that meets the real-time requirements of the system. The output of the command \texttt{uname -a} can be observed in code~\ref{output:uname_a_virt}.

@@

\section{Salamander 4 Bare Metal}\label{sec:salamander4-bare-metal}
The output of the command \texttt{uname -a} for Salamander 4 bare metal is shown in code~\ref{output:uname_a_hw}.

@@

\noindent As a reference point, the latency program was executed on Salamander 4 bare metal for a duration of 10 minutes. The complete command used was \texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of min, avg, max latencies in a Gnuplot-compatible format to the file \texttt{gnuplot.txt}. Figure~\ref{fig:max_latency_hardware} shows the gathered latency values in microseconds.

@@

\noindent With the help of the script depicted in code~\ref{script:qemu_def}, Salamander 4 is started together with the necessary hardware components in the QEMU environment. This makes it possible to run Salamander 4 on a variety of host systems, regardless of the specific hardware of the host. The following is a description of the components used for the virtualization of Salamander 4:
\begin{itemize}
	\item \textbf{bzImage}: Compressed Linux kernel image that is loaded by QEMU at system start. ``bz`` stands for big-zipped.
	\item \textbf{drive-c}: Directory serving as C drive for QEMU system, created and filled by qemu\_def.sh script.
	\item \textbf{ovmf.code.qcow2}: Firmware file for QEMU that enables UEFI boot process. \texttt{OVMF} stands for Open Virtual Machine Firmware, \texttt{qcow2} is a format for disk image files used by QEMU, it stands for "QEMU Copy On Write version 2".
	\item \textbf{qemu\_def.sh}: Shell script that starts QEMU with required parameters to boot Salamamder 4 OS. It is described in Code~\ref{script:qemu_def}.
	\item \textbf{salamander-image-sigmatek-core2.ext4}: Disk image of the Salamander 4 OS for the Sigmatek Core 2 platform. It uses the \texttt{ext4} file system and serves as the root file system in the QEMU virtual machine, acting as the virtual hard drive.
	\item \textbf{stek-drive-c-image-sigmatek-core2.tar.gz}: Compressed tarball containing a pre-configured environment for the Salamander 4 OS. It is unpacked and sets up the \texttt{drive-c/} directory with system and log files in the \texttt{qemu\_def.sh} script. 
	\item \textbf{vmlinux}: Uncompressed Linux kernel image, typically used for debugging.
\end{itemize}

@@

\noindent This script is run on a generic Ubuntu 22.04.4 system, as mentioned previously in chapter~\ref{cha:methodology}. The kernel version and other details are presented in Code~\ref{output:uname_a_ub}, using the \texttt{uname -a} command.

@@

\section{BIOS Configurations}

@@

\section{Kernel Configurations}

@@

	\item \textbf{isolcpus=4}: Isolates CPU 4 from the general scheduler, meaning no process will be scheduled to run on this CPU unless it is explicitly assigned. CPU isolation is explained in detail in section \ref{sec:cpu_isolation}.

@@

	\item \textbf{irqaffinity=0}: Sets the default Interrupt Request affinity to none. This means that no CPU core is preferred over another for handling IRQs. Instead, it lets the operating system decide how to distribute these IRQs across all the CPUs. Section \ref{sec:irq_handling} dives deeper into Interrupt Request affinity.

@@

\section{Host OS Configurations}
\subsection{CPU isolation}\label{sec:cpu_isolation}

Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved. Also, modifying the \texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process, as described later in section~\ref{sec:irq_handling}.

\noindent For CPU isolation, the \texttt{isolcpus} function was used to isolate a performance CPU from the general scheduling algorithms of the operating system. This means that the isolated CPUs will not be used for regular task scheduling, allowing them to be dedicated for the real-time specific tasks or processes. However, the isolcpus function only isolates at the user level and does not affect kernel tasks. Consequently, these kernel tasks and interrupts can still utilize the CPU~\cite{maPerformanceTuningKVMbased}. Output~\ref{output:pse} shows the user and kernel tasks that run on CPU 4. After the isolation, user tasks other than the QEMU process have been removed from running on this CPU. Only few critical kernel threads that are tied to this CPU still take CPU time.

\vspace{1em}
\begin{minipage}{0.95\columnwidth}
	\begin{lstlisting}[name={User and Kernel Tasks},label={output:pse},escapeinside={(*@}{@*)}]
		sigma_ibo@sigma-ibo:~$ cat /sys/devices/system/cpu/isolated
		4
		sigma_ibo@sigma-ibo:~$ ps axHo psr,pid,lwp,args,policy,nice,rtprio | awk '$1 == 4'
		4      38      38 [cpuhp/4]                   TS    0      -
		4      39      39 [idle_inject/4]             FF    -     50
		4      40      40 [migration/4]               FF    -     99
		4      41      41 [ksoftirqd/4]               TS    0      -
		4      43      43 [kworker/4:0H-events_highpr TS  -20      -
		4     505     505 [irq/189-iwlwifi:queue_1]   FF    -     50
		4     519     519 [irq/203-iwlwifi:exception] FF    -     50
		4    3008    3008 [kworker/4:1H-kblockd]      TS  -20      -
		4  177779  177779 [kworker/4:2-events]        TS    0      -
		4  448807  448807 [kworker/4:1-events]        TS    0      -		
\end{lstlisting}
\end{minipage}
\subsection{Interrupt Requests Handling}\label{sec:irq_handling}
Once the CPUs were isolated, Interrupt Requests handling was the next step. Interrupt Requests are used to send a signal to the CPU, prompting it to `interrupt' its current task and divert its attention to another task. This allows hardware devices to communicate with the CPU through frequent context switches, which can lead to performance degradation, especially in high-performance computing or real-time scenarios. To mitigate this, the IRQs needed to be removed from the isolated CPU. This was done by manipulating a file in the \texttt{proc} filesystem, namely \texttt{/proc/irq/<IRQ>/smp\_affinity}. The value in the \texttt{smp\_affinity} file is a bitmask in hexadecimal format. Each bit in this mask corresponds to a CPU in the system. The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the significance increases towards the left until CPU19. In a system with 14 CPUs, if every CPU was reserved for one IRQ, the value for \texttt{smp\_affinity} would be 3FFF. The script in code~\ref{script:smp_affinity} was written to check and log the distribution of Interrupt Requests across each CPU in Salamander 4.

\vspace{1em}
\begin{minipage}{0.95\columnwidth}
	\begin{lstlisting}[name={Check distribution of Interrupt Requests across each CPU},label={script:smp_affinity}]
		#!/bin/bash
		# Check if a command-line argument is provided
		if [ -z "$1" ]; then
			echo "Please provide a CPU number as a command-line argument."
			exit 1
		fi
		# Get the CPU number from the command-line argument
		CPU=$1
		# Initialize an empty array to store the IRQ numbers
		IRQs=()
		for IRQ in /proc/irq/*; do
			if [ -f "$IRQ/smp_affinity" ]; then
				# Read the current smp_affinity
				AFFINITY=$(cat "$IRQ/smp_affinity")
				# Check if the bit for the current CPU is set
				if (( (0x$AFFINITY & (1 << CPU)) != 0 )); then
					# Add the IRQ number to the array
					IRQs+=("${IRQ#/proc/irq/}")
				fi
			fi
		done
		# Sort the array
		IFS=$'\n' sorted=($(sort -n <<<"${IRQs[*]}"))
		# Print the CPU number
		echo "CPU $CPU IRQ affinity:"
		# Print the sorted IRQ numbers on separate lines
		for irq in "${sorted[@]}"; do
			echo "$irq"
		done

\end{lstlisting}
\end{minipage}

\noindent By changing the values of the \texttt{smp\_affinity} files of the respective IRQs, the assignment of IRQs was controlled so that they would not be handled by the isolated CPU. This reduced the interruptions caused by IRQs and the isolated CPUs were able to focus more on their assigned tasks.
\subsection{Disable dynamic frequency scaling}
\subsection{Disable RT throttling}
\subsection{No unexpected RT processes are running on your system}
\subsection{IRQ affinity}
\subsection{RCU CPU offloading}
\subsection{Suppress rcu cpu stall}
\subsection{Maybe?}
\subsection{Start QEMU normally and give all QEMU threads rt-priority} 
\subsection{Kill all running user processes}
\subsection{Set CPU Affinity for systemd services}
\subsection{Cache Isolation for CPU and GPU}
\subsection{Set CPU Affinity of IRQ thread to CPU 0}
\subsection{Set Device Driver Work Queue to CPU 0}
\subsection{Disable Machine Check}
\subsection{Stop Certain Services}
\section{QEMU-KVM Configurations}
\subsection{Tune lapic timer advance}
\subsection{Set QEMU options for real-time VM}
\subsection{Set CPU affinity and scheduling policy of QEMU CPU threads}
\subsection{Passthrough PCI devices into the VM}

@@

\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host. However, the isolcpus function only isolates at the user level and does not affect kernel tasks. Consequently, these kernel tasks and interrupts can still utilize the CPU. This led to the investigation of the causes for the observed high and inconsistent latency. The guest operates within the kvm\_entry and \texttt{kvm\_exit} events of the host. Kernelshark revealed a high frequency of \texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host. This frequent switching hinders the guest\'s ability to run continuously, thereby increasing the virtualization latency. To further understand this, trace-cmd was employed to trace various events in the host-guest communication, including the reasons for these events. Specifically, the causes for \texttt{kvm\_exit} events were analyzed. The command sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all was executed on the host for a duration of 5 seconds. The results in Figure~\ref{fig:kvm_exit_taskset} were obtained. Additionally, Table~\ref{tab:kvm_exit} provides a short description of the observed \texttt{kvm\_exit} events.

@@

\clearpage
Figure~\ref{fig:kvm_exit_taskset} shows \texttt{kvm\_exit} frequency with CPU islation.
\begin{figure}[H]
	\centering
	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_taskset.png}
	\caption[kvm exits]{kvm exits}
	\label{fig:kvm_exit_taskset}
\end{figure}

Figure~\ref{fig:kvm_exits_default} shows \texttt{kvm\_exit} frequency without CPU islation.
\begin{figure}[H]
	\centering
	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_default.png}
	\caption[kvm exits default]{kvm exits default}
	\label{fig:kvm_exits_default}
\end{figure}

@@

	\noindent Without CPU isolation, context switches take place at operating system level and not at hypervisor level. This explains why there are fewer \texttt{kvm\_exit} events. However, this will also, as previously shown, lead to higher latency, as context switches at operating system level generally take longer than a \texttt{kvm\_exit} and kvm\_entry.

	\noindent When the CPU was dedicated to the QEMU process, on the other hand,there was a significant increase in \texttt{kvm\_exit} events. This is because every context switch takes place at hypervisor level. Nevertheless, lower latency was achieved thereby, as the qemu process is no longer influenced by the CPU scheduling of the operating system.

@@

	\noindent In the process of analyzing the \texttt{kvm\_exit} events, several reasons for these exits were identified. The most frequent among these were the \texttt{APIC\_WRITE} and \texttt{HLT} events. The former is initiated when the guest writes to its Advanced Programmable Interrupt Controller (APIC), a component of the CPU that manages hardware interrupts. The latter occurs when the guest executes the HLT instruction, effectively halting the CPU until the next external interrupt is fired. Other significant but less frequent events included \texttt{EXTERNAL\_INTERRUPT} and \texttt{IO\_INSTRUCTION}. These events are indicative of the guest's interaction with hardware devices and its execution of I/O operations. Events such as \texttt{EPT\_MISCONFIG} and \texttt{PREEMPTION\_TIMER} were also noted. These could potentially signal issues with memory management and the host's scheduling of the guest. While events like \texttt{PAUSE\_INSTRUCTION}, \texttt{EPT\_VIOLATION}, \texttt{EOI\_INDUCED}, \texttt{MSR\_READ}, and \texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.

@@

Figure~\ref{fig:kvm_exit_vapic} shows \texttt{kvm\_exit} frequency with APIC virtualization. Comparing this to the previous Figures~\ref{fig:kvm_exit_taskset} and~\ref{fig:kvm_exits_default}, it can be observed that an APIC\_Write no longer occurs.

>>>>

\noindent Figure~\ref{fig:lstopo} is the output of the ~\texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices. 
\begin{figure}[H]
	\centering
	\includegraphics[width=1.0\columnwidth]{img/lstopo.png}
	\caption[Hardware topology]{Hardware topology}
	\label{fig:lstopo}
\end{figure}

@@

\section{Structure}
\noindent Salamander 4 is the proprietary operating system of SIGMATEK. It is based on Linux version 5.15.94 and integrates Xenomai 3.2, a real-time development environment~\cite{XenomaiXenomai}. Salamander 4 is a 64-bit system, which refers to the x86\_64 architecture. The real-time behaviour is achieved through the use of Symmetric Multi-Processing (SMP) and Preemptive Scheduling (PREEMPT). In addition, it uses IRQPIPE to process interrupts in a way that meets the real-time requirements of the system. The output of the command ~\texttt{uname -a} can be observed in code~\ref{output:uname_a_virt}.

@@

\section{Salamander 4 Bare Metal}\label{sec:salamander4-bare-metal}
The output of the command ~\texttt{uname -a} for Salamander 4 bare metal is shown in code~\ref{output:uname_a_hw}.

@@

\noindent As a reference point, the latency program was executed on Salamander 4 bare metal for a duration of 10 minutes. The complete command used was ~\texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of min, avg, max latencies in a Gnuplot-compatible format to the file ~\texttt{gnuplot.txt}. Figure~\ref{fig:max_latency_hardware} shows the gathered latency values in microseconds.

@@

\noindent With the help of the script depicted in code~\ref{script:qemu_def}, Salamander 4 is started together with the necessary hardware components in the QEMU environment. This makes it possible to run Salamander 4 on a variety of host systems, regardless of the specific hardware of the host. The following is a description of the components used for the virtualization of Salamander 4:
\begin{itemize}
	\item \textbf{bzImage}: Compressed Linux kernel image that is loaded by QEMU at system start. ``bz`` stands for big-zipped.
	\item \textbf{drive-c}: Directory serving as C drive for QEMU system, created and filled by qemu\_def.sh script.
	\item \textbf{ovmf.code.qcow2}: Firmware file for QEMU that enables UEFI boot process. ~\texttt{OVMF} stands for Open Virtual Machine Firmware, ~\texttt{qcow2} is a format for disk image files used by QEMU, it stands for "QEMU Copy On Write version 2".
	\item \textbf{qemu\_def.sh}: Shell script that starts QEMU with required parameters to boot Salamamder 4 OS. It is described in Code~\ref{script:qemu_def}.
	\item \textbf{salamander-image-sigmatek-core2.ext4}: Disk image of the Salamander 4 OS for the Sigmatek Core 2 platform. It uses the ~\texttt{ext4} file system and serves as the root file system in the QEMU virtual machine, acting as the virtual hard drive.
	\item \textbf{stek-drive-c-image-sigmatek-core2.tar.gz}: Compressed tarball containing a pre-configured environment for the Salamander 4 OS. It is unpacked and sets up the ~\texttt{drive-c/} directory with system and log files in the ~\texttt{qemu\_def.sh} script. 
	\item \textbf{vmlinux}: Uncompressed Linux kernel image, typically used for debugging.
\end{itemize}

@@

\noindent This script is run on a generic Ubuntu 22.04.4 system, as mentioned previously in chapter~\ref{cha:methodology}. The kernel version and other details are presented in Code~\ref{output:uname_a_ub}, using the ~\texttt{uname -a} command.

@@

\section{BIOS Configurations}\label{sec:bios_configurations}

@@

\section{Kernel Configurations}\label{sec:kernel_configurations}

@@

	\item \textbf{isolcpus=4}: Isolates CPU 4 from the general scheduler, meaning no process will be scheduled to run on this CPU unless it is explicitly assigned. CPU isolation is explained in detail in section \ref{subsec:cpu_isolation}.

@@

	\item \textbf{irqaffinity=0}: Sets the default Interrupt Request affinity to none. This means that no CPU core is preferred over another for handling IRQs. Instead, it lets the operating system decide how to distribute these IRQs across all the CPUs. Section \ref{subsec:irq_handling} dives deeper into Interrupt Request affinity.

@@

\section{Host OS Configurations}\label{sec:host_configurations}
\subsection{CPU isolation}\label{subsec:cpu_isolation}

Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved. Also, modifying the ~\texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process, as described later in section~\ref{subsec:irq_handling}.

\noindent For CPU isolation, the ~\texttt{isolcpus} function was used to isolate a performance CPU from the general scheduling algorithms of the operating system. This means that the isolated CPUs will not be used for regular task scheduling, allowing them to be dedicated for the real-time specific tasks or processes. However, the isolcpus function only isolates at the user level and does not affect kernel tasks. Consequently, these kernel tasks and interrupts can still utilize the CPU~\cite{maPerformanceTuningKVMbased}. Output~\ref{output:pse} shows the user and kernel tasks that run on CPU 4. After the isolation, user tasks other than the QEMU process have been removed from running on this CPU. Only few per-CPU kernel threads that are tied to this CPU still take CPU time.

\vspace{1em}
\begin{minipage}{0.95\columnwidth}
	\begin{lstlisting}[name={User and Kernel Tasks},label={output:pse},escapeinside={(*@}{@*)}]
		sigma_ibo@sigma-ibo:~$ cat /sys/devices/system/cpu/isolated
		4
		sigma_ibo@sigma-ibo:~$ ps axHo psr,pid,lwp,args,policy,nice,rtprio | awk '$1 == 4'
		4      38      38 [cpuhp/4]                   TS    0      -
		4      39      39 [idle_inject/4]             FF    -     50
		4      40      40 [migration/4]               FF    -     99
		4      41      41 [ksoftirqd/4]               TS    0      -
		4      42      42 [kworker/4:0-events]        TS    0      -
		4      43      43 [kworker/4:0H-kblockd]      TS  -20      -
		4     153     153 [kworker/4:1-events]        TS    0      -
		4   81649   81649 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   81654 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   81676 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   81702 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   82185 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   82187 qemu-system-x86_64 -M pc,ac TS    0      -
		4   82134   82134 [kworker/4:1H-kblockd]      TS  -20      -
\end{lstlisting}
\end{minipage}
\subsection{Interrupt Affinity}\label{subsec:irq_handling}
Once the CPUs were isolated, interrupt requests handling was the next step. The purpose of interrupt requests is to inform the CPU to stop working on a certain job and start working on another. This allows hardware devices to communicate with the CPU through frequent context switches, which can introduce latency in real-time systems. The ~\texttt{/proc/interrupts} file can be monitored using the command ~\texttt{watch -d -c cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time. The IRQs needed to be removed from the isolated CPU by manipulating the ~\texttt{/proc/irq/<IRQ>/smp\_affinity} files. The value in the ~\texttt{smp\_affinity} file is a bitmask in hexadecimal format where each bit corresponds to a CPU. The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the most significant bit (MSB) on the left corresponds to the last CPU (CPU13). When there are 14 CPUs available, the default value for ~\texttt{smp\_affinity} would be 3FFF. Removing CPU4 out of this bitmask would be setting bit five to zero, resulting in 3FEF (11 1111 1110 1111). The python script in Code~\ref{script:smp_affinity} was written to show a table of the distribution of interrupt requests across each CPU. By changing the values of the ~\texttt{smp\_affinity} files, the assignment of IRQs to CPUs is controlled. However, as the proc filesystem resets to its default state after each reboot, manually changing numerous IRQ files is tedious and time-consuming. This process was automated using the shell script in Code~\ref{script:change_smp_affinity}, which is executed after every reboot.

\vspace{1em}
\begin{minipage}{0.95\columnwidth}
	\begin{lstlisting}[name={Check distribution of interrupt requests across each CPU},label={script:smp_affinity}]
		import os
		import pandas as pd
		from tabulate import tabulate

		# Get the number of CPUs
		num_cpus = os.cpu_count()

		# Initialize a dictionary to store the CPUs for each IRQ
		irqs = {}

		# Iterate over each IRQ
		for irq in os.listdir('/proc/irq'):
			# Check if the smp_affinity file exists for this IRQ
			if os.path.isfile(f'/proc/irq/{irq}/smp_affinity'):
				# Read the current smp_affinity
				with open(f'/proc/irq/{irq}/smp_affinity', 'r') as f:
					affinity = int(f.read().strip(), 16)
				# Initialize an empty list to store the CPUs for this IRQ
				cpus = []
				# Iterate over each CPU
				for cpu in range(num_cpus):
					# Check if the bit for the current CPU is set
					if ((affinity & (1 << cpu)) != 0):
						# Add the CPU to the list for this IRQ
						cpus.append(cpu)
				# Sort the list of CPUs
				cpus.sort()
				# Add the list of CPUs to the dictionary for this IRQ
				irqs[irq] = cpus

		# Create a DataFrame to store the table
		df = pd.DataFrame(index=sorted(irqs.keys(), key=int), columns=range(num_cpus))

		# Fill the DataFrame with 'x' where a CPU is assigned to an IRQ
		for irq, cpus in irqs.items():
			for cpu in cpus:
				df.loc[irq, cpu] = 'x'

		# Replace NaN values with empty strings
		df.fillna('', inplace=True)

		# Print the table in pipe format
		print(tabulate(df, headers='keys', tablefmt='pipe', showindex=True))

		# Convert the DataFrame to a markdown table
		markdown_table = df.to_markdown()

		# Write the markdown table to a file
		with open('table_CPU_IRQ.md', 'w') as f:
			f.write(markdown_table)
\end{lstlisting}
\end{minipage}

\vspace{1em}
\begin{minipage}{0.95\columnwidth}
	\begin{lstlisting}[name={Change IRQ assignment of a CPU},label={script:change_smp_affinity}]
		#!/bin/bash

		# Check if a command-line argument is provided
		if [ -z "$1" ]; then
			echo "Please provide a CPU number as a command-line argument."
			exit 1
		fi

		# Get the CPU number from the command-line argument
		CPU=$1

		# Define the mask values
		declare -A mask_values
		mask_values=( [0]="3ffe" [1]="3ffd" [2]="3ffb" [3]="3ff7" [4]="3fef" [5]="3fdf" [6]="3fbf" [7]="3f7f" [8]="3eff" [9]="3dff" [10]="3bff" [11]="37ff" [12]="2fff" [13]="17ff")

		# Run the check_smp_affinity.sh script and get the IRQs
		IRQs=$(./check_smp_affinity.sh $CPU | grep -o '[0-9]\+')

		# Initialize an empty array to store the IRQs that could not be removed
		failed_IRQs=()
		# Initialize an empty array to store the IRQs that were successfully removed
		succeeded_IRQs=()

		# Loop over the IRQs
		for IRQ in $IRQs; do
			# Try to change the smp_affinity
			echo ${mask_values[$CPU]} | sudo tee /proc/irq/$IRQ/smp_affinity > /dev/null 2>&1
			# If the command failed, add the IRQ to the failed_IRQs array
			if [ $? -ne 0 ]; then
				failed_IRQs+=($IRQ)
			else
				succeeded_IRQs+=($IRQ)
			fi
		done

		# Check if there were any failed IRQs
		if [ ${#failed_IRQs[@]} -ne 0 ]; then
			echo "IRQs ${failed_IRQs[@]} could not be removed from CPU $CPU."
		fi

		# Check if there were any successful IRQs
		if [ ${#succeeded_IRQs[@]} -ne 0 ]; then
			# Remove the first entry from the succeeded_IRQs array
			succeeded_IRQs=("${succeeded_IRQs[@]:1}")
			echo "IRQs ${succeeded_IRQs[@]} were removed from CPU $CPU."
		fi		
\end{lstlisting}
\end{minipage}
\subsection{Disable RT throttling}
If a real-time task consumes 100\% of the CPU time, the system may become unresponsive as a whole. This happens because an RT process is constantly using the CPU and the Linux scheduler will not schedule other non-RT processes in the meantime. To prevent complete system lockups, the kernel has a function to throttle RT processes if they consume 0.95 seconds out of every 1 second of CPU time. It does this by pausing the process for the remaining 0.05 seconds, which is not desired because this could result in missed deadlines. RT throttling can be disabled by  writing the value -1 to the ~\texttt{/proc/sys/kernel/sched\_rt\_runtime\_us} file. This change also needs to be made permanent because the proc filesystem resets to its default state after each reboot. For this purpose, the line ~\texttt{kernel.sched\_rt\_runtime\_us = -1} can be appended to the end of the ~\texttt{/etc/sysctl.conf} file, which is read at boot time and used to configure kernel parameters. This reduces the potential for missed deadlines. 
\subsection{No unexpected RT processes}
It is important that there are no other unexpected real time processes running on the system. A high priority RT process could occasionally be created by the host operating system at boot, hence disabling the offending RT process is advised if it is not needed. The process of disabling these RT processes is usually described in the documentation for the chosen Linux distribution.
\subsection{Suppress rcu cpu stall}
As already mentioned in section \ref{sec:kernel_configurations}, the Linux kernel uses RCU as a synchronization mechanism for reading from and writing to shared data. A stall occurs when an RCU read-side critical section runs for an extended period of time, preventing the RCU grace period from completing. By writing 1 to the \texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_suppress} file, these warnings are suppressed. This helps eliminate potential interruptions and CPU load that are caused by these warnings, thereby reduce the potential for increased latency. This setting is also not persistent across reboots, so the command needs to be added to a startup script. 
\subsection{Maybe?}

\subsection{Start QEMU normally and give all QEMU threads rt-priority}

\subsection{Cache Isolation for CPU and GPU}

\subsection{Set CPU Affinity of IRQ thread to CPU 0}

\subsection{Set Device Driver Work Queue to CPU 0}

\subsection{Set CPU Affinity for systemd services}

\subsection{Stop Certain Services}

\subsection{Disable Machine Check}

There are other configurations that may be relevant depending on your use case, some of which are documented in [this talk](https://www.youtube.com/watch?v=NrjXEaTSyrw) and [this other talk](https://www.youtube.com/watch?v=w3yT8zJe0Uw). Additionally, quality-of-life configurations, such the variables in /etc/security/limits.conf, may need to be tuned as well.

\section{QEMU-KVM Configurations}\label{sec:guest_configurations}
\subsection{Tune lapic timer advance}
\subsection{Set QEMU options for real-time VM}
\subsection{Set CPU affinity and scheduling policy of QEMU CPU threads}
\subsection{Passthrough PCI devices into the VM}

@@

\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host. However, the isolcpus function only isolates at the user level and does not affect kernel tasks. Consequently, these kernel tasks and interrupts can still utilize the CPU. This led to the investigation of the causes for the observed high and inconsistent latency. The guest operates within the kvm\_entry and ~\texttt{kvm\_exit} events of the host. Kernelshark revealed a high frequency of ~\texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host. This frequent switching hinders the guest\'s ability to run continuously, thereby increasing the virtualization latency. To further understand this, trace-cmd was employed to trace various events in the host-guest communication, including the reasons for these events. Specifically, the causes for ~\texttt{kvm\_exit} events were analyzed. The command sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all was executed on the host for a duration of 5 seconds. The results in Figure~\ref{fig:kvm_exit_taskset} were obtained. Additionally, Table~\ref{tab:kvm_exit} provides a short description of the observed ~\texttt{kvm\_exit} events.

@@

\clearpage
Figure~\ref{fig:kvm_exit_taskset} shows ~\texttt{kvm\_exit} frequency with CPU islation.
\begin{figure}[H]
	\centering
	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_taskset.png}
	\caption[kvm exits]{kvm exits}
	\label{fig:kvm_exit_taskset}
\end{figure}

Figure~\ref{fig:kvm_exits_default} shows ~\texttt{kvm\_exit} frequency without CPU islation.
\begin{figure}[H]
	\centering
	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_default.png}
	\caption[kvm exits default]{kvm exits default}
	\label{fig:kvm_exits_default}
\end{figure}

@@

	\noindent Without CPU isolation, context switches take place at operating system level and not at hypervisor level. This explains why there are fewer ~\texttt{kvm\_exit} events. However, this will also, as previously shown, lead to higher latency, as context switches at operating system level generally take longer than a ~\texttt{kvm\_exit} and kvm\_entry.

	\noindent When the CPU was dedicated to the QEMU process, on the other hand,there was a significant increase in ~\texttt{kvm\_exit} events. This is because every context switch takes place at hypervisor level. Nevertheless, lower latency was achieved thereby, as the qemu process is no longer influenced by the CPU scheduling of the operating system.

@@

	\noindent In the process of analyzing the ~\texttt{kvm\_exit} events, several reasons for these exits were identified. The most frequent among these were the ~\texttt{APIC\_WRITE} and ~\texttt{HLT} events. The former is initiated when the guest writes to its Advanced Programmable Interrupt Controller (APIC), a component of the CPU that manages hardware interrupts. The latter occurs when the guest executes the HLT instruction, effectively halting the CPU until the next external interrupt is fired. Other significant but less frequent events included ~\texttt{EXTERNAL\_INTERRUPT} and ~\texttt{IO\_INSTRUCTION}. These events are indicative of the guest's interaction with hardware devices and its execution of I/O operations. Events such as ~\texttt{EPT\_MISCONFIG} and ~\texttt{PREEMPTION\_TIMER} were also noted. These could potentially signal issues with memory management and the host's scheduling of the guest. While events like ~\texttt{PAUSE\_INSTRUCTION}, ~\texttt{EPT\_VIOLATION}, ~\texttt{EOI\_INDUCED}, ~\texttt{MSR\_READ}, and ~\texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.

@@

Figure~\ref{fig:kvm_exit_vapic} shows ~\texttt{kvm\_exit} frequency with APIC virtualization. Comparing this to the previous Figures~\ref{fig:kvm_exit_taskset} and~\ref{fig:kvm_exits_default}, it can be observed that an APIC\_Write no longer occurs.
